Propositions as Sessions, Mechanically

### 2015-04-21 paper/presentation.tex
Some boilerplate for slides.
Working on the slide content for presentation.

\usepackage[latin1]{inputenc}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{amsmath}
\usepackage{listings}
\usepackage{comment}
\usepackage{tikz}
\usetikzlibrary{arrows,positioning}
\usepackage[style=alphabetic,natbib=true]{biblatex}

@@

@@

today i am going to talk about a formalisation effort based on an existing
type system whose properties have previously only been studied
informally. first in order to motivate the previous and current work i shall
make two general observations here. one: distributed communication systems are
becoming increasingly prevalent in modern society. two: the correct
functioning of these systems hinges on the protocol between any two
communicating agents to be well-defined and correctly implemented.

\begin{frame}
Session types by example
\end{frame}

one approach to achieve such correctness guarantees is by using binary session
types. [INTRODUCE BINARY SESSION TYPES HERE]. specifying a communication
protocol then involves providing mutual dual types to the communicating
parties. [a simple example with two processes is displayed].

@@

there has been a lot of activity in this area producing a variety of different
session-based type systems. many systems build on top of previous work adding
more complex features such as asynchronous, polymorphic and recursive session
types. extending informal proofs is error-prone and difficult because it is
not always clear which parts of a system's metatheory is affected by a
change. a formalised system would reduce likelihood of errors in the proofs
and aid extension since the theorem prover would highlight changes to the
metatheory

@@

the formalisation i present here is an encoding of a system in the Coq proof
assistant. it is based on a two-tier system providing a logical foundation for
understanding session types. A high-level overview of the system is as
follows: we have a process calculus CP with operators from classical linear
logic, a high-level functional language GV with session types, and a
translation from GV to CP which interprets operators in classical linear logic
as session types.

\begin{frame}
an example of CP classical linear logic operators and their session type
interpretation
\end{frame}

by classical linear logic i mean a logic where negation is an involution which
restricts the use of certain rules on resources, namely that all linear
resources must be used (corresponding to prohibiting weakening) and that one
cannot duplicate a linear resource (corresponding to contraction). some
operators in the CP calculus are presented as they appear in the Coq encoding
(note that they are very similar to their informal counterparts). the
interpretation of tensor product is output of A and continue as B. The dual of
this is input of A and continue as the negation of B. Thus negation in
classical linear logic corresponds to session duality.

\begin{frame}
an example of GV and a translation???
\end{frame}

GV has built-in support for session types. consider the rule for sending a
value of type T along a channel N of type !T.S. the translation to CP is
presented. blah blah something about the session types being a little odd or
overlook this probably

@@

the system has three main properties. subject reduction of CP, top-level
cut-elimination and translation preserving well-typedness. One out of the
three has been formalised and some issues and guidelines have been developed
accordingly. some comments on progress towards the other two is offered

The system enjoys freedom from deadlock cut elimination in CP corresponds to
communication in GV.

subject reduction is formalised using a relation on processes

cut elimination is not currently proven since we require a termination measure
on the length of a derivation and a relationship between the subprocesses of a
cut. other work based on formalising logics where the derivations are explicit
may be inspirational

the translation seems to require a duplication of the specifications for cp
and gv typing rules not done

@@

here are some issues and things learnt! -- our guidelines are all about
modularisation, representation, library reuse, automation!

\begin{frame}
\frametitle{Preamble}
\item Propositions as Sessions
\item system features
\item Coq
\item encoding
\item subject reduction theorem
\item difficulties in formalisation
\item implications
\item future work
\end{frame}

>>>>

\usepackage[latin1]{inputenc}
\usepackage{lstmacros}
\usepackage{graphicx}
\usepackage{array}
\usepackage{subfigure}
\usepackage{amsmath}
\usepackage{comment}
\usepackage{etoolbox}
\usepackage{tikz}
\usetikzlibrary{arrows,positioning}
\usepackage[style=alphabetic,natbib=true]{biblatex}

@@

\newtoggle{tdlp}
\toggletrue{tdlp}
\input{propmacros}
\newcommand{\altt}{~|~}

@@

\begin{comment}
today i am going to talk about a formalisation effort based on an existing
type system whose properties have previously only been studied
informally. first in order to motivate the previous and current work i shall
make two general observations here. one: distributed communication systems are
becoming increasingly prevalent in modern society. two: the correct
functioning of these systems hinges on the protocol between any two
communicating agents to be well-defined and correctly implemented.
\end{comment}

\begin{frame}
\frametitle{Session Types}
\begin{math}
  \begin{array}{rclc}
    S & \bnf  & \inpt{T}{S} & \mbox{input value of type T, continue as S}\\
      & \altt & \outpt{T}{S} & \mbox{output value of type T, continue as S}\\
      & \altt & \Branch{l_i}{S_i}{i\in I} &
                           \mbox{offer between $i$ alternatives}\\
      & \altt & \Choice{l_i}{S_i}{i\in I} &
                           \mbox{choose between $i$ alternatives}\\
      & \altt & \mbox{\lstinline{end}} & \mbox{terminate session}
 \end{array}
\end{math}
\end{frame}

\begin{comment}
one approach to achieve such correctness guarantees is by using binary session
types. [INTRODUCE BINARY SESSION TYPES HERE]. specifying a communication
protocol then involves providing mutual dual types to the communicating
parties. [a simple example with two processes is displayed].
\end{comment}

@@

\begin{comment}
there has been a lot of activity in this area producing a variety of different
session-based type systems. many systems build on top of previous work adding
more complex features such as asynchronous, polymorphic and recursive session
types. extending informal proofs is error-prone and difficult because it is
not always clear which parts of a system's metatheory is affected by a
change. a formalised system would reduce likelihood of errors in the proofs
and aid extension since the theorem prover would highlight changes to the
metatheory
\end{comment}

@@

\begin{comment}
the formalisation i present here is an encoding of a system in the Coq proof
assistant. it is based on a two-tier system providing a logical foundation for
understanding session types. A high-level overview of the system is as
follows: we have a process calculus CP with operators from classical linear
logic, a high-level functional language GV with session types, and a
translation from GV to CP which interprets operators in classical linear logic
as session types.
\end{comment}

\begin{frame}
negation as involution
an example of CP classical linear logic operators and their session type
interpretation (times/with)
\end{frame}

\begin{comment}
by classical linear logic i mean a logic where negation is an involution which
restricts the use of certain rules on resources, namely that all linear
resources must be used (corresponding to prohibiting weakening) and that one
cannot duplicate a linear resource (corresponding to contraction). some typing
rules for the CP calculus are presented as they appear in the Coq encoding
(note that they are very similar to their informal counterparts). the
interpretation of tensor product is output of A and continue as B. The dual of
this is input of A and continue as the negation of B. Thus negation in
classical linear logic corresponds to session duality.
\end{comment}

\begin{frame}
an example of GV
\end{frame}

\begin{comment}
GV has built-in support for session types. consider the rule for sending a
value of type T along a channel N of type !T.S.
\end{comment}

\begin{frame}
something about a translation
\end{frame}

\begin{comment}
the translation to CP is
presented. blah blah something about the session types being a little odd or
overlook this probably
\end{comment}

@@

\begin{comment}
the system has three main properties. subject reduction of CP, top-level
cut-elimination and translation preserving well-typedness. of which subject
reduction is the only one fully formalised. the other two presented some
challenges but some comments on progress towards the other two is
offered. i also highlight some issues and guidelines
\end{comment}

\begin{frame}
subject reduction
\end{frame}

\begin{comment}
subject reduction is formalised using a relation on processes
\end{comment}

\begin{frame}
problems for cut elimination
\end{frame}

\begin{comment}
sessions correspond to propositions
processes correspond to proofs
cut elimination corresponds to communication

cut elimination is not currently proven since we require a termination measure
on the length of a derivation and a relationship between the subprocesses of a
cut. other work based on formalising logics where the derivations are explicit
may be inspirational
\end{comment}

\begin{frame}
problems of the translation
\end{frame}

\begin{comment}
the translation seems to require a duplication of the specifications for cp
and gv typing rules not done

@@

here are some issues and things learnt! -- our guidelines are all about
modularisation, representation, library reuse, automation!
\end{comment}

\begin{frame}
\frametitle{Preamble}
\begin{itemize}
\item Propositions as Sessions
\item system features
\item Coq
\item encoding
\item subject reduction theorem
\item difficulties in formalisation
\item implications
\item future work
\end{itemize}
\end{frame}
